---
title: "Using a simplified a, v model"
author:
- name: "Paul L. Maurizio"
  affiliation: 
  - The University of Chicago, Department of Medicine, Chicago, IL, USA
  - The University of North Carolina at Chapel Hill, Department of Genetics, Chapel Hill, NC, USA (previous)

date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Using a simplified a, v model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Load packages

```{r, message=FALSE}
library(MCMCglmm)
library(litterDiallel)
library(xtable)
```

## Load data, add column

```{r, message=FALSE}
data("litters")
litters$inbred <- ifelse(litters$Dam_Founder == litters$Sire_Founder, 1, 0)
```

## Generate design matrices

```{r, message=FALSE}
matrices <- diallelMatrixMaker_av(data=litters, dam.col.name="Dam_Founder", sire.col.name="Sire_Founder",
                                  batch.col.name="YearMonth", batch.1.col.name="litterorder")
strains <- c("AJ", "B6", "129", "NOD", "NZO", "CAST", "PWK", "WSB")
matrices.Mt <- diallelMatrixMakeAndRotate_av(litters, dam.col.name="Dam_Founder", 
  sire.col.name="Sire_Founder", batch.col.name="YearMonth",
  batch.1.col.name="litterorder", n.strains=length(strains))
M.matrices <- matrices.Mt[[1]]
t.matrices <- matrices.Mt[[2]]
```

## Define design matrices for each random effects class

```{r, message=FALSE}
add.mat <- matrices$add.mat
jk.mat <- matrices$jk.mat
batch.mat <- matrices$batch.mat
batch.1.mat <- matrices$batch.1.mat
```

## Define reduced-dimension design matrices to pre-center variable estimates in each class
### See: Lenarcic et al., 2012, *Genetics*; and Crowley et al., 2014, *Genetics*

```{r, message=FALSE}
t.add.mat <- t.matrices$t.add.mat
t.jk.mat <- t.matrices$t.jk.mat
t.batch.mat <- t.matrices$t.batch.mat
t.batch.1.mat <- t.matrices$t.batch.1.mat
```

## Define parameters to model the data as approximately Gaussian, after using a variance-stabilizing transform.

```{r, message=FALSE}
nitt <- 1515
burnin <- 15
thin <- 1
# nitt <- 15150
# burnin <- 150
# thin <- 5
dist <- "gaussian"
ordernum <- 0
adjust <- 0
litters$SqrtWeaned <- sqrt(litters$Weaned)
```

## Define priors

```{r, message=FALSE}
priors <- list(B=list(V=1e+03*diag(2), mu=c(0,0)),
              R = list(V = diag(1), nu = 0.002),
              G = list(
                  G1=list(V=1, nu=0.002), 
                  G2=list(V=1, nu=0.002), 
                  G3=list(V=1, nu=0.002), 
                  G4=list(V=1, nu=0.002)))
```

## Fit the model

```{r, message=FALSE}
fit1 <- MCMCglmm(SqrtWeaned ~ 1 + litternum,
                 random = ~ idv(t.add.mat) +
                   idv(t.jk.mat) + 
                   idv(t.batch.mat) + idv(t.batch.1.mat),
                 prior=priors,
                 nitt=nitt, burnin = burnin,
                 thin=thin, pr=TRUE,
                 family = "gaussian", data=litters)
```

## Retrieve MCMC chains

```{r, message=FALSE}
allChains <- as.mcmc(cbind(fit1$Sol,fit1$VCV))
allChains.fixed <- allChains[,c("(Intercept)", "litternum")]
colnames(allChains.fixed) <- c("mu", "litternum")
allChains.add <- allChains[, grep(colnames(allChains), pattern="^t.add.mat[^.]", value=TRUE)]
allChains.jk <- allChains[, grep(colnames(allChains), pattern="^t.jk.mat[^.]", value=TRUE)]
allChains.batch <- allChains[, grep(colnames(allChains), pattern="^t.batch.mat[^.]", value=TRUE)]
allChains.batch.1 <- allChains[, grep(colnames(allChains), pattern="^t.batch.1.mat[^.]", value=TRUE)]
allChains.sigma <- allChains[, c('t.add.mat.', 't.jk.mat.', 
                                 't.batch.mat.', 't.batch.1.mat.', 'units')]
```

## Rotate parameter estimates back to full parameter space

```{r, message=FALSE}
M.add <- M.matrices$M.add
M.jk <- M.matrices$M.jk
M.batch <- M.matrices$M.batch
M.batch.1 <- M.matrices$M.batch.1

allChains.add <- allChains.add %*% t(M.add)
allChains.jk <- allChains.jk %*% t(M.jk)
allChains.batch <- allChains.batch %*% t(M.batch)
allChains.batch.1 <- allChains.batch.1 %*% t(M.batch.1)

colnames(allChains.add) <- colnames(add.mat)
colnames(allChains.jk) <- colnames(jk.mat)
colnames(allChains.batch) <- colnames(batch.mat)
colnames(allChains.batch.1) <- colnames(batch.1.mat)
```

## Edit parameter class names, combine chains

```{r, message=FALSE}
colnames(allChains.sigma) <- c("tau-sq-add",
                               "tau-sq-jk",
                               "tau-sq-year-month",
                               "tau-sq-litterorder",
                               "sigma-sq")

allChains <- as.mcmc(cbind(allChains.fixed, allChains.add, 
                   allChains.jk, 
                   allChains.batch, allChains.batch.1,
                   allChains.sigma))
```

## Plot MCMC traces

```{r, message=FALSE}
#plot(allChains[,1])
```

## Plot diallel effect estimates

```{r, message=FALSE, fig_width=3, fig_height=8}
xlim=c(-0.3, 0.3);

var.labels.all <- varnames(allChains)
var.labels <- colnames(add.mat)
plot_hpd(allChains[,var.labels], main="general effects", xlim=xlim); abline(v=0, col="grey")

var.labels <- colnames(jk.mat)
plot_hpd(allChains[,var.labels], main="strainpair-specific", xlim=xlim); abline(v=0, col="grey")
```

## Plot covariate effect estimates

```{r, message=FALSE, fig_width=3, fig_height=8}
var.labels.all <- varnames(allChains)
var.labels <- c(colnames(batch.mat))[1:24]
plot_hpd(allChains[,var.labels], main="wean date", xlim=xlim); abline(v=0, col="grey")
var.labels <- c(colnames(batch.mat))[25:48]
plot_hpd(allChains[,var.labels], main="wean date", xlim=xlim); abline(v=0, col="grey")
var.labels <- c("litternum", colnames(batch.1.mat))
plot_hpd(allChains[,var.labels], main="litter order", xlim=xlim); abline(v=0, col="grey")
```

## Print tables

```{r, results="asis"}
#hpd.table1 <- summary(allChains)[[1]]
hpd.table2 <- summary(allChains)[[2]]
#print(xtable(hpd.table1), type="html")
print(xtable(hpd.table2, digits=4), type="html")
```


## Plot VarComps

```{r, messages=FALSE}
allChains.sigma.names <- c("tau-sq-add",
                           "tau-sq-jk",
                           "sigma-sq")
allChains.sigma <- allChains[,allChains.sigma.names]

var.df <- cbind(as.data.frame(allChains[, allChains.sigma.names]), adjust)
names(var.df)[names(var.df)=="var1"] <- "adjust"
var.df$total <- var.df$`tau-sq-add` +
                var.df$`tau-sq-jk` +
                var.df$`sigma-sq` + var.df$`adjust`
varcomp <- NULL
varcomp$additive <- var.df$`tau-sq-add` / var.df$`total`
varcomp$symmetric.epistatic <- var.df$`tau-sq-jk` / var.df$`total`

varcomp$total.explained <- (var.df$`total` - var.df$`sigma-sq` - var.df$`adjust`)/ var.df$`total`
varcomp$noise <- (var.df$`sigma-sq` + var.df$`adjust`)/var.df$`total`
varcomp <- as.data.frame(varcomp)

varcomp.table <- cbind(mean=colMeans(varcomp), 
	lower.bound=HPDinterval(as.mcmc(varcomp))[,1],
	upper.bound=HPDinterval(as.mcmc(varcomp))[,2])

# varcolors.y0 <- c(7.5, 6.5, 5.5, 4.5, 3.5)
# varcolors.y1 <- c(6.5, 5.5, 4.5, 3.5, 2.5)
psq <- data.frame(varcomp.table)

par(mar = c(5.1, 13.5, 2.1, 12.1))
rows <- nrow(psq)
plot(y = c(0.5, rows + 0.5), x = c(-0.06, 1.01), yaxs = "i",
    xaxs = "i", col = "white", ylab = "", xlab = "variance explained",
    yaxt = "n")
abline(v = c(0, 0.2, 0.4, 0.6, 0.8, 1), lty = 2, col = "grey")
abline(h = 2.5, lwd = 1.5)
axis(side = 2, at = c(rows:1), labels = rownames(psq),
    las = 1)
rightlabels <- paste(as.character(sprintf("%.2f", round(100 *
    psq$mean, digits = 2))), "% (", as.character(sprintf("%.2f",
    round(100 * psq$lower.bound, digits = 2))), "%, ", as.character(sprintf("%.2f",
    round(100 * psq$upper.bound, digits = 2))), "%)", sep = "")
axis(side = 4, at = c(rows:1), labels = rightlabels, tick = FALSE,
    las = 1)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "black",
    cex = 1.5)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "white",
    cex = 0.5)
```

## Table of results

```{r, results="asis"}
print(xtable(psq, digits=4), type="html")
```

## Plot VarPs

```{r, messages=FALSE}
allChains.df <- data.frame(allChains, check.names=FALSE)

allChains.fixed <- allChains.df[,c("mu", "litternum")]
allChains.add <- as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^additive[^.]", value=TRUE)])
allChains.jk <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^v[^.]", value=TRUE)])
allChains.batch <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^YearMonth[^.]", value=TRUE)])
allChains.batch.1 <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^order[^.]", value=TRUE)])
allChains.sigma <-   as.matrix(cbind(allChains.df[, grep(colnames(allChains.df), pattern="^tau-sq-[^.]", value=TRUE)],
                      `sigma-sq`=allChains.df$`sigma-sq`))

allChains <- as.mcmc(cbind(allChains.fixed, allChains.add, allChains.jk,
                           allChains.batch, allChains.batch.1,
                           allChains.sigma))

founder.names <- c("AJ", "B6", "129", "NOD", "NZO", "CAST", "PWK", "WSB")
projection <- as.data.frame(as.matrix(expand.grid(founder.names, founder.names)[,c(2,1)]))

colnames(projection) <- c("Dam_Founder", "Sire_Founder")
projection.matrices <- diallelMatrixMaker_av(projection, dam.col.name="Dam_Founder", sire.col.name="Sire_Founder")

proj.add <- apply(allChains.add, MARGIN=1, FUN=function(x){projection.matrices$add.mat %*% x})
proj.jk <- apply(allChains.jk, MARGIN=1, FUN=function(x){projection.matrices$jk.mat %*% x})
proj.mu <- matrix(rep(allChains.fixed[,"mu"],each=dim(projection)[1]),nrow=dim(projection)[1])

proj.epsilon <- sapply(allChains.sigma[,"sigma-sq"], FUN=function(x){rnorm(n=dim(projection)[1], mean=0, sd=sqrt(x))})
proj.determ.mu <- proj.mu + proj.add + proj.jk 
proj.determ <- proj.add + proj.jk

proj.not.add <- proj.determ - proj.add
proj.not.jk <- proj.determ - proj.jk
proj.total <- proj.determ + proj.epsilon
proj.y <- proj.determ.mu + proj.epsilon

ss.add <- NULL
ss.jk <- NULL
ss.epsilon <- NULL
ss.denom <- NULL

for(i in 1:dim(proj.add)[2]){
  ss.add[i] <- sum((proj.determ[,i] - proj.not.add[,i])^2)
  ss.jk[i] <- sum((proj.determ[,i] - proj.not.jk[,i])^2)
  ss.epsilon[i] <- sum((proj.y[,i] - proj.determ.mu[,i])^2)
}

Y.bar <- matrix(rep(apply(proj.determ, MARGIN=2, FUN=sum)/dim(projection)[1], each=dim(projection)[1]), nrow=dim(projection)[1])
ss.denom <- ss.add + ss.jk + ss.epsilon + adjust

additive <- ss.add/ss.denom
epistatic.symm <- ss.jk/ss.denom
explained <- (ss.add + ss.jk)/ss.denom
unexplained <- 1 - explained

varps.all <- cbind(additive, symmetric.epistatic=epistatic.symm, total.explained=explained, noise=unexplained)
varps.all.table <- cbind(mean=colMeans(varps.all), 
	lower.bound=HPDinterval(as.mcmc(varps.all))[,1],
	upper.bound=HPDinterval(as.mcmc(varps.all))[,2])

psq <- data.frame(varps.all.table)

par(mar = c(5.1, 13.5, 2.1, 12.1))
rows <- nrow(psq)
plot(y = c(0.5, rows + 0.5), x = c(-0.06, 1.01), yaxs = "i",
    xaxs = "i", col = "white", ylab = "", xlab = "Variance Projection",
    yaxt = "n")
abline(v = c(0, 0.2, 0.4, 0.6, 0.8, 1), lty = 2, col = "grey")
abline(h = 2.5, lwd = 1.5)
segments(x0 = psq$lower.bound, x1 = psq$upper.bound, y0 = c(rows:1),
    y1 = c(rows:1), lwd = 4, lend=1)
axis(side = 2, at = c(rows:1), labels = rownames(psq),
    las = 1)
rightlabels <- paste(as.character(sprintf("%.2f", round(100 *
    psq$mean, digits = 2))), "% (", as.character(sprintf("%.2f",
    round(100 * psq$lower.bound, digits = 2))), "%, ", as.character(sprintf("%.2f",
    round(100 * psq$upper.bound, digits = 2))), "%)", sep = "")
axis(side = 4, at = c(rows:1), labels = rightlabels, tick = FALSE,
    las = 1)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "black",
    cex = 1.5)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "white",
    cex = 0.5)
```

## Table of results

```{r, results="asis"}
print(xtable(psq, digits=4), type="html")
```
