---
title: "Analyzing the litter diallel"
author:
- name: "Paul L. Maurizio"
  affiliation: 
  - The University of Chicago, Department of Medicine, Chicago, IL, USA
  - The University of North Carolina at Chapel Hill, Department of Genetics, Chapel Hill, NC, USA (previous)

date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Analyzing the litter diallel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Load packages

```{r, message=FALSE}
library("MCMCglmm")
library("litterDiallel")
library("PLMcctools")
library("xtable")
```

## Load data, add column

```{r, message=FALSE}
data("litters")
litters$inbred <- ifelse(litters$Dam_Founder == litters$Sire_Founder, 1, 0)
```

## Generate design matrices

```{r, message=FALSE}
matrices <- diallelMatrixMaker(data=litters, dam.col.name="Dam_Founder", sire.col.name="Sire_Founder", 
                               batch.col.name="YearMonth", batch.1.col.name="litterorder")
M.matrices <- diallelMatrixMakeAndRotate(litters, dam.col.name="Dam_Founder", 
  sire.col.name="Sire_Founder", batch.col.name="YearMonth",
  batch.1.col.name="litterorder")[[1]]
t.matrices <- diallelMatrixMakeAndRotate(litters, dam.col.name="Dam_Founder", 
  sire.col.name="Sire_Founder", batch.col.name="YearMonth",
  batch.1.col.name="litterorder")[[2]]
```

## Define design matrices for each random effects class

```{r, message=FALSE}
add.mat <- matrices$add.mat
mat.mat <- matrices$mat.mat
dam.mat <- matrices$dam.mat
sire.mat <- matrices$sire.mat
inbred.mat <- matrices$inbred.mat
jk.mat <- matrices$jk.mat
asymm.mat <- matrices$asymm.mat
batch.mat <- matrices$batch.mat
batch.1.mat <- matrices$batch.1.mat
```

## Define reduced-dimension design matrices to pre-center variable estimates in each class
### See: Lenarcic et al., 2012, *Genetics*; and Crowley et al., 2014, *Genetics*

```{r, message=FALSE}
t.add.mat <- t.matrices$t.add.mat
t.mat.mat <- t.matrices$t.mat.mat
t.dam.mat <- t.matrices$t.dam.mat
t.sire.mat <- t.matrices$t.sire.mat
t.inbred.mat <- t.matrices$t.inbred.mat
t.jk.mat <- t.matrices$t.jk.mat
t.asymm.mat <- t.matrices$t.asymm.mat
t.batch.mat <- t.matrices$t.batch.mat
t.batch.1.mat <- t.matrices$t.batch.1.mat
```

## Define parameters to model the data as approximately Gaussian, after using a variance-stabilizing transform.

```{r, message=FALSE}
nitt <- 1515
burnin <- 15
thin <- 1
# nitt <- 15150
# burnin <- 150
# thin <- 5
dist <- "gaussian"
ordernum <- 0
adjust <- 0
litters$SqrtWeaned <- sqrt(litters$Weaned)
```

## Define priors

```{r, message=FALSE}
priors <- list(B=list(V=1e+03*diag(3), mu=c(0,0,0)),
              R = list(V = diag(1), nu = 0.002),
              G = list(
                  G1=list(V=1, nu=0.002), 
                  G2=list(V=1, nu=0.002), 
                  G3=list(V=1, nu=0.002), 
                  G4=list(V=1, nu=0.002), 
                  G5=list(V=1, nu=0.002), 
                  G6=list(V=1, nu=0.002),
                  G7=list(V=1, nu=0.002)))
```

## Fit the model

```{r, message=FALSE}
fit1 <- MCMCglmm(SqrtWeaned ~ 1 + inbred + litternum,
                 random = ~ idv(t.add.mat) + idv(t.mat.mat) + idv(t.inbred.mat) +
                   idv(t.jk.mat) + idv(t.asymm.mat) +
                   idv(t.batch.mat) + idv(t.batch.1.mat),
                 prior=priors,
                 nitt=nitt, burnin = burnin,
                 thin=thin, pr=TRUE,
                 family = "gaussian", data=litters)
```

## Retrieve MCMC chains

```{r, message=FALSE}
allChains <- as.mcmc(cbind(fit1$Sol,fit1$VCV))
allChains.fixed <- allChains[,c("(Intercept)", "inbred", "litternum")]
colnames(allChains.fixed) <- c("mu", "inbred", "litternum")
allChains.add <- allChains[, grep(colnames(allChains), pattern="^t.add.mat[^.]", value=TRUE)]
allChains.mat <- allChains[, grep(colnames(allChains), pattern="^t.mat.mat[^.]", value=TRUE)]
allChains.inbred <- allChains[, grep(colnames(allChains), pattern="^t.inbred.mat[^.]", value=TRUE)]
allChains.jk <- allChains[, grep(colnames(allChains), pattern="^t.jk.mat[^.]", value=TRUE)]
allChains.asymm <- allChains[, grep(colnames(allChains), pattern="^t.asymm.mat[^.]", value=TRUE)]
allChains.batch <- allChains[, grep(colnames(allChains), pattern="^t.batch.mat[^.]", value=TRUE)]
allChains.batch.1 <- allChains[, grep(colnames(allChains), pattern="^t.batch.1.mat[^.]", value=TRUE)]
allChains.sigma <- allChains[, c('t.add.mat.', 't.mat.mat.', 't.inbred.mat.', 
                                 't.jk.mat.', 't.asymm.mat.', 
                                 't.batch.mat.', 't.batch.1.mat.', 'units')]
```

## Rotate parameter estimates back to full parameter space

```{r, message=FALSE}
M.add <- M.matrices$M.add
M.mat <- M.matrices$M.mat
M.inbred <- M.matrices$M.inbred
M.jk <- M.matrices$M.jk
M.asymm <- M.matrices$M.asymm
M.batch <- M.matrices$M.batch
M.batch.1 <- M.matrices$M.batch.1

allChains.add <- allChains.add %*% t(M.add)
allChains.mat <- allChains.mat %*% t(M.mat)
allChains.inbred <- allChains.inbred %*% t(M.inbred)
allChains.jk <- allChains.jk %*% t(M.jk)
allChains.asymm <- allChains.asymm %*% t(M.asymm)
allChains.batch <- allChains.batch %*% t(M.batch)
allChains.batch.1 <- allChains.batch.1 %*% t(M.batch.1)

colnames(allChains.add) <- colnames(add.mat)
colnames(allChains.mat) <- colnames(mat.mat)
colnames(allChains.inbred) <- colnames(inbred.mat)
colnames(allChains.jk) <- colnames(jk.mat)
colnames(allChains.asymm) <- colnames(asymm.mat)
colnames(allChains.batch) <- colnames(batch.mat)
colnames(allChains.batch.1) <- colnames(batch.1.mat)
```

## Edit parameter class names, combine chains

```{r, message=FALSE}
colnames(allChains.sigma) <- c("tau-sq-add",
                               "tau-sq-maternal",
                               "tau-sq-inbred",
                               "tau-sq-jk",
                               "tau-sq-asymm",
                               "tau-sq-year-month",
                               "tau-sq-litterorder",
                               "sigma-sq")

allChains <- as.mcmc(cbind(allChains.fixed, allChains.add, 
                   allChains.mat, allChains.inbred,
                   allChains.jk, allChains.asymm,
                   allChains.batch, allChains.batch.1,
                   allChains.sigma))
```

## Plot MCMC traces

```{r, message=FALSE}
#plot(allChains[,1])
```

## Plot diallel effect estimates

```{r, message=FALSE, fig_width=3, fig_height=8}
jk.ordering <- c('v:B6;AJ', 'v:129;AJ', 'v:NOD;AJ', 'v:NZO;AJ', 'v:CAST;AJ', 'v:PWK;AJ', 'v:WSB;AJ',
                 'v:129;B6', 'v:NOD;B6', 'v:NZO;B6', 'v:CAST;B6', 'v:PWK;B6', 'v:WSB;B6',
                 'v:NOD;129', 'v:NZO;129', 'v:CAST;129', 'v:PWK;129', 'v:WSB;129',
                 'v:NZO;NOD', 'v:CAST;NOD', 'v:PWK;NOD', 'v:WSB;NOD',
                 'v:CAST;NZO', 'v:PWK;NZO', 'v:WSB;NZO', 'v:PWK;CAST', 'v:WSB;CAST', 'v:WSB;PWK')

asymm.ordering <- c('w:B6;AJ', 'w:129;AJ', 'w:NOD;AJ', 'w:NZO;AJ', 'w:CAST;AJ', 'w:PWK;AJ', 'w:WSB;AJ',
                 'w:129;B6', 'w:NOD;B6', 'w:NZO;B6', 'w:CAST;B6', 'w:PWK;B6', 'w:WSB;B6',
                 'w:NOD;129', 'w:NZO;129', 'w:CAST;129', 'w:PWK;129', 'w:WSB;129',
                 'w:NZO;NOD', 'w:CAST;NOD', 'w:PWK;NOD', 'w:WSB;NOD',
                 'w:CAST;NZO', 'w:PWK;NZO', 'w:WSB;NZO', 'w:PWK;CAST', 'w:WSB;CAST', 'w:WSB;PWK')

varcolors <- c("#A6CEE3", "#B2DF8A", "#FB9A99",
               "#FDBF6F", "#CAB2D6", "#D2B48C")
varcolors.y0 <- rev(c(0.25, 8.25, 9.25, 17.25))
varcolors.y1 <- rev(c(7.75, 8.75, 16.75, 24.75))
xlim=c(-0.3, 0.3);

var.labels.all <- varnames(allChains)
var.labels <- c(colnames(add.mat), colnames(mat.mat), "inbred", colnames(inbred.mat))
plot.hpd(allChains[,var.labels], main="general effects", xlim=xlim); abline(v=0, col="grey")
segments(x0=xlim[1]*0.85, x1=xlim[1]*0.85, y0=varcolors.y0, y1=varcolors.y1, col=varcolors[1:4], lwd=7, lend=1)

var.labels <- jk.ordering
plot.hpd(allChains[,var.labels], main="strainpair-specific", xlim=xlim); abline(v=0, col="grey")
segments(x0=xlim[1]*0.85, x1=xlim[1]*0.85, y0=0.25, y1=27.75, col=varcolors[5], lwd=7, lend=1)

var.labels <- asymm.ordering
plot.hpd(allChains[,var.labels], main="strainpair-specific", xlim=xlim); abline(v=0, col="grey")
segments(x0=xlim[1]*0.85, x1=xlim[1]*0.85, y0=0.25, y1=27.75, col=varcolors[6], lwd=7, lend=1)

```

## Plot covariate effect estimates

```{r, message=FALSE, fig_width=3, fig_height=8}
var.labels.all <- varnames(allChains)
var.labels <- c(colnames(batch.mat))[1:24]
plot.hpd(allChains[,var.labels], main="wean date", xlim=xlim); abline(v=0, col="grey")
var.labels <- c(colnames(batch.mat))[25:48]
plot.hpd(allChains[,var.labels], main="wean date", xlim=xlim); abline(v=0, col="grey")
var.labels <- c("litternum", colnames(batch.1.mat))
plot.hpd(allChains[,var.labels], main="litter order", xlim=xlim); abline(v=0, col="grey")
```

## Print tables

```{r, results="asis"}
#hpd.table1 <- summary(allChains)[[1]]
hpd.table2 <- summary(allChains)[[2]]
#print(xtable(hpd.table1), type="html")
print(xtable(hpd.table2, digits=4), type="html")
```

## Calculate dam\_strain and sire\_strain effects

```{r, message=FALSE}
allChains.df <- as.data.frame(as.matrix(allChains))
allChains.df$`dam:AJ` <- allChains.df[,"additive:AJ"] + allChains.df[, "maternal:AJ"]
allChains.df$`dam:B6` <- allChains.df[,"additive:B6"] + allChains.df[, "maternal:B6"]
allChains.df$`dam:129` <- allChains.df[,"additive:129"] + allChains.df[, "maternal:129"]
allChains.df$`dam:NOD` <- allChains.df[,"additive:NOD"] + allChains.df[, "maternal:NOD"]
allChains.df$`dam:NZO` <- allChains.df[,"additive:NZO"] + allChains.df[, "maternal:NZO"]
allChains.df$`dam:CAST` <- allChains.df[,"additive:CAST"] + allChains.df[, "maternal:CAST"]
allChains.df$`dam:PWK` <- allChains.df[,"additive:PWK"] + allChains.df[, "maternal:PWK"]
allChains.df$`dam:WSB` <- allChains.df[,"additive:WSB"] + allChains.df[, "maternal:WSB"]

allChains.df$`sire:AJ` <- allChains.df[,"additive:AJ"] - allChains.df[, "maternal:AJ"]
allChains.df$`sire:B6` <- allChains.df[,"additive:B6"] - allChains.df[, "maternal:B6"]
allChains.df$`sire:129` <- allChains.df[,"additive:129"] - allChains.df[, "maternal:129"]
allChains.df$`sire:NOD` <- allChains.df[,"additive:NOD"] - allChains.df[, "maternal:NOD"]
allChains.df$`sire:NZO` <- allChains.df[,"additive:NZO"] - allChains.df[, "maternal:NZO"]
allChains.df$`sire:CAST` <- allChains.df[,"additive:CAST"] - allChains.df[, "maternal:CAST"]
allChains.df$`sire:PWK` <- allChains.df[,"additive:PWK"] - allChains.df[, "maternal:PWK"]
allChains.df$`sire:WSB` <- allChains.df[,"additive:WSB"] - allChains.df[, "maternal:WSB"]

allChains.dam <- allChains.df[, grep(names(allChains.df), pattern="^dam:", value=TRUE)]
allChains.sire <- allChains.df[, grep(names(allChains.df), pattern="^sire:", value=TRUE)]

allChains.bup <- allChains
allChains <- as.mcmc(cbind(allChains.fixed, allChains.add, allChains.mat, 
                           allChains.dam, allChains.sire, allChains.inbred,
                           allChains.jk, allChains.asymm,
                           allChains.batch, allChains.batch.1,
                           allChains.sigma))
```

## Plot dam\_strain and \sire_strain MCMC traces

```{r, messages=FALSE}
#plot(allChains[,colnames(dam.mat)]); plot(allChains[,colnames(sire.mat)])
```

## Plot dam\_strain and sire\_strain effects
```{r, messages=FALSE}
xlim=c(-0.4, 0.4);
var.labels.all <- varnames(allChains)
var.labels.new <- c("dam:AJ", "dam:B6", "dam:129", "dam:NOD", "dam:NZO", "dam:CAST", "dam:PWK", "dam:WSB",
                "sire:AJ", "sire:B6", "sire:129", "sire:NOD", "sire:NZO", "sire:CAST", "sire:PWK", "sire:WSB")
var.labels <- c(colnames(dam.mat), colnames(sire.mat))#, "inbred", colnames(inbred.mat))
plot.hpd(allChains[,var.labels[1:8]], xlim=xlim, main="dam strain", main.line=1.75); abline(v=0, col="grey")
plot.hpd(allChains[,var.labels[9:16]], xlim=xlim, main="sire strain", main.line=1.75); abline(v=0, col="grey")
```

## Plot VarComps

```{r, messages=FALSE}
allChains.sigma.names <- c("tau-sq-add",
                           "tau-sq-maternal",
                           "tau-sq-inbred",
                           "tau-sq-jk",
                           "tau-sq-asymm",
                           "sigma-sq")
allChains.sigma <- allChains[,allChains.sigma.names]

var.df <- cbind(as.data.frame(allChains[, allChains.sigma.names]), adjust)
names(var.df)[names(var.df)=="var1"] <- "adjust"
var.df$total <- var.df$`tau-sq-add` + var.df$`tau-sq-maternal` + var.df$`tau-sq-inbred` +
                var.df$`tau-sq-jk` + var.df$`tau-sq-asymm` + 
                var.df$`sigma-sq` + var.df$`adjust`
varcomp <- NULL
varcomp$additive <- var.df$`tau-sq-add` / var.df$`total`
varcomp$parental.sex <- var.df$`tau-sq-maternal` / var.df$`total`
varcomp$inbred <- var.df$`tau-sq-inbred` / var.df$`total`
varcomp$symmetric.epistatic <- var.df$`tau-sq-jk` / var.df$`total`
varcomp$asymmetric.epistatic <- var.df$`tau-sq-asymm` / var.df$`total`

varcomp$total.explained <- (var.df$`total` - var.df$`sigma-sq` - var.df$`adjust`)/ var.df$`total`
varcomp$noise <- (var.df$`sigma-sq` + var.df$`adjust`)/var.df$`total`
varcomp <- as.data.frame(varcomp)

varcomp.table <- cbind(mean=colMeans(varcomp), 
	lower.bound=HPDinterval(as.mcmc(varcomp))[,1],
	upper.bound=HPDinterval(as.mcmc(varcomp))[,2])

varcolors <- c("#A6CEE3", "#B2DF8A", #"#FB9A99",
               "#FDBF6F", "#CAB2D6", "#D2B48C")
varcolors.y0 <- c(7.5, 6.5, 5.5, 4.5, 3.5)
varcolors.y1 <- c(6.5, 5.5, 4.5, 3.5, 2.5)
psq <- data.frame(varcomp.table)

par(mar = c(5.1, 13.5, 2.1, 12.1))
rows <- nrow(psq)
plot(y = c(0.5, rows + 0.5), x = c(-0.06, 1.01), yaxs = "i",
    xaxs = "i", col = "white", ylab = "", xlab = "variance explained",
    yaxt = "n")
segments(x0=-0.05, x1=-0.05, y0=varcolors.y0, y1=varcolors.y1, col=varcolors, lwd=15, lend=1)
abline(v = c(0, 0.2, 0.4, 0.6, 0.8, 1), lty = 2, col = "grey")
abline(h = 2.5, lwd = 1.5)
segments(x0 = psq$lower.bound, x1 = psq$upper.bound, y0 = c(rows:1),
    y1 = c(rows:1), lwd = 4, lend=1)
axis(side = 2, at = c(rows:1), labels = rownames(psq),
    las = 1)
rightlabels <- paste(as.character(sprintf("%.2f", round(100 *
    psq$mean, digits = 2))), "% (", as.character(sprintf("%.2f",
    round(100 * psq$lower.bound, digits = 2))), "%, ", as.character(sprintf("%.2f",
    round(100 * psq$upper.bound, digits = 2))), "%)", sep = "")
axis(side = 4, at = c(rows:1), labels = rightlabels, tick = FALSE,
    las = 1)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "black",
    cex = 1.5)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "white",
    cex = 0.5)
```

## Table of results

```{r, results="asis"}
print(xtable(psq, digits=4), type="html")
```

## Plot VarPs

```{r, messages=FALSE}
allChains.df <- data.frame(allChains, check.names=FALSE)

allChains.df$`dam:AJ` <- allChains.df[,"additive:AJ"] + allChains.df[, "maternal:AJ"]
allChains.df$`dam:B6` <- allChains.df[,"additive:B6"] + allChains.df[, "maternal:B6"]
allChains.df$`dam:129` <- allChains.df[,"additive:129"] + allChains.df[, "maternal:129"]
allChains.df$`dam:NOD` <- allChains.df[,"additive:NOD"] + allChains.df[, "maternal:NOD"]
allChains.df$`dam:NZO` <- allChains.df[,"additive:NZO"] + allChains.df[, "maternal:NZO"]
allChains.df$`dam:CAST` <- allChains.df[,"additive:CAST"] + allChains.df[, "maternal:CAST"]
allChains.df$`dam:PWK` <- allChains.df[,"additive:PWK"] + allChains.df[, "maternal:PWK"]
allChains.df$`dam:WSB` <- allChains.df[,"additive:WSB"] + allChains.df[, "maternal:WSB"]

allChains.df$`sire:AJ` <- allChains.df[,"additive:AJ"] - allChains.df[, "maternal:AJ"]
allChains.df$`sire:B6` <- allChains.df[,"additive:B6"] - allChains.df[, "maternal:B6"]
allChains.df$`sire:129` <- allChains.df[,"additive:129"] - allChains.df[, "maternal:129"]
allChains.df$`sire:NOD` <- allChains.df[,"additive:NOD"] - allChains.df[, "maternal:NOD"]
allChains.df$`sire:NZO` <- allChains.df[,"additive:NZO"] - allChains.df[, "maternal:NZO"]
allChains.df$`sire:CAST` <- allChains.df[,"additive:CAST"] - allChains.df[, "maternal:CAST"]
allChains.df$`sire:PWK` <- allChains.df[,"additive:PWK"] - allChains.df[, "maternal:PWK"]
allChains.df$`sire:WSB` <- allChains.df[,"additive:WSB"] - allChains.df[, "maternal:WSB"]

allChains.dam <- allChains.df[, grep(names(allChains.df), pattern="^dam:", value=TRUE)]
allChains.sire <- allChains.df[, grep(names(allChains.df), pattern="^sire:", value=TRUE)]

allChains.fixed <- allChains.df[,c("mu", "inbred", "litternum")]
allChains.add <- as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^additive[^.]", value=TRUE)])
allChains.mat <-  as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^maternal[^.]", value=TRUE)])
allChains.dam <-  as.matrix( allChains.df[, grep(colnames(allChains.df), pattern="^dam[^.]", value=TRUE)])
allChains.sire <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^sire[^.]", value=TRUE)])
allChains.inbred <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^inbred[^.]", value=TRUE)])
allChains.jk <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^v[^.]", value=TRUE)])
allChains.asymm <-  as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^w[^.]", value=TRUE)])
allChains.batch <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^YearMonth[^.]", value=TRUE)])
allChains.batch.1 <-   as.matrix(allChains.df[, grep(colnames(allChains.df), pattern="^order[^.]", value=TRUE)])
allChains.sigma <-   as.matrix(cbind(allChains.df[, grep(colnames(allChains.df), pattern="^tau-sq-[^.]", value=TRUE)],
                      `sigma-sq`=allChains.df$`sigma-sq`))

allChains <- as.mcmc(cbind(allChains.fixed, allChains.add, allChains.mat, 
                           allChains.dam, allChains.sire, allChains.inbred,
                           allChains.jk, allChains.asymm,
                           allChains.batch, allChains.batch.1,
                           allChains.sigma))

founder.names <- c("AJ", "B6", "129", "NOD", "NZO", "CAST", "PWK", "WSB")
projection <- as.data.frame(as.matrix(expand.grid(founder.names, founder.names)[,c(2,1)]))

projection$inbred <- ifelse(projection[,1]==projection[,2], 1, 0)
projection.inbred.mat <- diag(projection$inbred)
colnames(projection) <- c("Dam_Founder", "Sire_Founder", "inbred")
projection.matrices <- diallelMatrixMaker(projection, dam.col.name="Dam_Founder", sire.col.name="Sire_Founder")

proj.add <- apply(allChains.add, MARGIN=1, FUN=function(x){projection.matrices$add.mat %*% x})
proj.mat <- apply(allChains.mat, MARGIN=1, FUN=function(x){projection.matrices$mat.mat %*% x})
proj.dam <- apply(allChains.dam, MARGIN=1, FUN=function(x){projection.matrices$dam.mat %*% x})
proj.sire <- apply(allChains.sire, MARGIN=1, FUN=function(x){projection.matrices$sire.mat %*% x})
proj.inbred <- apply(allChains.inbred, MARGIN=1, FUN=function(x){projection.matrices$inbred.mat %*% x})
proj.jk <- apply(allChains.jk, MARGIN=1, FUN=function(x){projection.matrices$jk.mat %*% x})
proj.asymm <- apply(allChains.asymm, MARGIN=1, FUN=function(x){projection.matrices$asymm.mat %*% x})
proj.mu <- matrix(rep(allChains.fixed[,"mu"],each=dim(projection)[1]),nrow=dim(projection)[1])
proj.inbred.overall <- projection.inbred.mat %*% matrix(rep(allChains.fixed[,"inbred"],each=dim(projection)[1]), nrow=dim(projection)[1])

proj.epsilon <- sapply(allChains.sigma[,"sigma-sq"], FUN=function(x){rnorm(n=dim(projection)[1], mean=0, sd=sqrt(x))})
proj.determ.mu <- proj.mu + proj.inbred.overall + proj.add + proj.mat + proj.inbred + proj.jk + proj.asymm
proj.determ <- proj.add + proj.mat + proj.inbred + proj.jk + proj.asymm
# The following projections should be equivalent:
# proj.determ.mu <- proj.mu + proj.inbred.overall + proj.dam + proj.sire + proj.inbred + proj.jk + proj.asymm
# proj.determ <- proj.dam + proj.sire + proj.inbred + proj.jk + proj.asymm
proj.not.add <- proj.determ - proj.add
proj.not.mat <- proj.determ - proj.mat
proj.not.dam <- proj.determ - proj.dam
proj.not.sire <- proj.determ - proj.sire
proj.not.inbred <- proj.determ - proj.inbred
proj.not.inbred.overall <- proj.determ - proj.inbred.overall
proj.not.jk <- proj.determ - proj.jk
proj.not.asymm <- proj.determ - proj.asymm
proj.total <- proj.determ + proj.epsilon
proj.y <- proj.determ.mu + proj.epsilon

ss.add <- NULL
ss.mat <- NULL
ss.dam <- NULL
ss.sire <- NULL
ss.inbred <- NULL
ss.inbred.overall <- NULL
ss.jk <- NULL
ss.asymm <- NULL
ss.epsilon <- NULL
ss.denom <- NULL

for(i in 1:dim(proj.add)[2]){
  ss.add[i] <- sum((proj.determ[,i] - proj.not.add[,i])^2)
  ss.mat[i] <- sum((proj.determ[,i] - proj.not.mat[,i])^2)
  ss.dam[i] <- sum((proj.determ[,i] - proj.not.dam[,i])^2)
  ss.sire[i] <- sum((proj.determ[,i] - proj.not.sire[,i])^2)
  ss.inbred[i] <- sum((proj.determ[,i] - proj.not.inbred[,i])^2)
  ss.inbred.overall[i] <- sum((proj.determ[,i] - proj.not.inbred.overall[,i])^2)
  ss.jk[i] <- sum((proj.determ[,i] - proj.not.jk[,i])^2)
  ss.asymm[i] <- sum((proj.determ[,i] - proj.not.asymm[,i])^2)
  ss.epsilon[i] <- sum((proj.y[,i] - proj.determ.mu[,i])^2)
}

Y.bar <- matrix(rep(apply(proj.determ, MARGIN=2, FUN=sum)/dim(projection)[1], each=dim(projection)[1]), nrow=dim(projection)[1])
ss.denom <- ss.add + ss.mat + ss.inbred + ss.inbred.overall + ss.jk + ss.asymm + ss.epsilon + adjust

additive <- ss.add/ss.denom
mat.dev <- ss.mat/ss.denom
maternal <- ss.dam/ss.denom
paternal <- ss.sire/ss.denom
inbred <- ss.inbred/ss.denom
inbred.overall <- ss.inbred.overall/ss.denom
epistatic.symm <- ss.jk/ss.denom
epistatic.asymm <- ss.asymm/ss.denom
explained <- (ss.add + ss.mat + ss.inbred + ss.jk + ss.asymm)/ss.denom
unexplained <- 1 - explained

varps.all <- cbind(additive, parental.sex=mat.dev, inbred, inbred.overall, symmetric.epistatic=epistatic.symm, asymmetric.epistatic=epistatic.asymm, total.explained=explained, noise=unexplained)
varps.all.table <- cbind(mean=colMeans(varps.all), 
	lower.bound=HPDinterval(as.mcmc(varps.all))[,1],
	upper.bound=HPDinterval(as.mcmc(varps.all))[,2])

varcolors <- c("#A6CEE3", "#B2DF8A", "#FB9A99",
               "#FDBF6F", "#CAB2D6", "#D2B48C")
varcolors.y0 <- c(8.5, 7.5, 6.5, 5.5, 4.5, 3.5)
varcolors.y1 <- c(7.5, 6.5, 5.5, 4.5, 3.5, 2.5)
psq <- data.frame(varps.all.table)

par(mar = c(5.1, 13.5, 2.1, 12.1))
rows <- nrow(psq)
plot(y = c(0.5, rows + 0.5), x = c(-0.06, 1.01), yaxs = "i",
    xaxs = "i", col = "white", ylab = "", xlab = "Variance Projection",
    yaxt = "n")
segments(x0=-0.05, x1=-0.05, y0=varcolors.y0, y1=varcolors.y1, col=varcolors, lwd=15, lend=1)
abline(v = c(0, 0.2, 0.4, 0.6, 0.8, 1), lty = 2, col = "grey")
abline(h = 2.5, lwd = 1.5)
segments(x0 = psq$lower.bound, x1 = psq$upper.bound, y0 = c(rows:1),
    y1 = c(rows:1), lwd = 4, lend=1)
axis(side = 2, at = c(rows:1), labels = rownames(psq),
    las = 1)
rightlabels <- paste(as.character(sprintf("%.2f", round(100 *
    psq$mean, digits = 2))), "% (", as.character(sprintf("%.2f",
    round(100 * psq$lower.bound, digits = 2))), "%, ", as.character(sprintf("%.2f",
    round(100 * psq$upper.bound, digits = 2))), "%)", sep = "")
axis(side = 4, at = c(rows:1), labels = rightlabels, tick = FALSE,
    las = 1)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "black",
    cex = 1.5)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "white",
    cex = 0.5)
```

## Table of results

```{r, results="asis"}
print(xtable(psq, digits=4), type="html")
```

## Plot dam\_strain and sire\_strain VarPs

```{r, messages=FALSE}
varps.all.damsire <- cbind(dam.strain=maternal, sire.strain=paternal, inbred, inbred.overall, symmetric.epistatic=epistatic.symm, asymmetric.epistatic=epistatic.asymm, total.explained=explained, noise=unexplained)

varps.all.damsire.table <- cbind(mean=colMeans(varps.all.damsire), 
  lower.bound=HPDinterval(as.mcmc(varps.all.damsire))[,1],
  upper.bound=HPDinterval(as.mcmc(varps.all.damsire))[,2])

## MCMCglmm model variance component estimates
varcolors <- c("#781112", "#131E3A", "#FB9A99",
               "#FDBF6F", "#CAB2D6", "#D2B48C")
varcolors.y0 <- c(8.5, 7.5, 6.5, 5.5, 4.5, 3.5)
varcolors.y1 <- c(7.5, 6.5, 5.5, 4.5, 3.5, 2.5)
psq <- data.frame(varps.all.damsire.table)

par(mar = c(5.1, 13.5, 2.1, 12.1))
rows <- nrow(psq)
plot(y = c(0.5, rows + 0.5), x = c(-0.06, 1.01), yaxs = "i",
    xaxs = "i", col = "white", ylab = "", xlab = "Variance Projection",
    yaxt = "n")
segments(x0=-0.05, x1=-0.05, y0=varcolors.y0, y1=varcolors.y1, col=varcolors, lwd=15, lend=1)
abline(v = c(0, 0.2, 0.4, 0.6, 0.8, 1), lty = 2, col = "grey")
abline(h = 2.5, lwd = 1.5)
segments(x0 = psq$lower.bound, x1 = psq$upper.bound, y0 = c(rows:1),
    y1 = c(rows:1), lwd = 4, lend=1)
axis(side = 2, at = c(rows:1), labels = rownames(psq),
    las = 1)
rightlabels <- paste(as.character(sprintf("%.2f", round(100 *
    psq$mean, digits = 2))), "% (", as.character(sprintf("%.2f",
    round(100 * psq$lower.bound, digits = 2))), "%, ", as.character(sprintf("%.2f",
    round(100 * psq$upper.bound, digits = 2))), "%)", sep = "")
axis(side = 4, at = c(rows:1), labels = rightlabels, tick = FALSE,
    las = 1)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "black",
    cex = 1.5)
points(x = psq$mean, y = c(rows:1), pch = 16, col = "white",
    cex = 0.5)
```

## Table of results

```{r, results="asis"}
print(xtable(psq, digits=4), type="html")
```

